---
title: "Prediction Assignment Write Up"
author: "Nica Stefan"
date: "30 octombrie 2017"
output: html_document
---

#1. Overview

The current document is the final report of Peer Assesment Project from Coursera's course Practical Machine Learning, specialization Data Science. It's build in RStudio, using its knitr functions and published in html format. The goal of the project is to predict how 6 participants performed some exercise as described below. This is the "classe" variable in the training set. The algorithm built here will be applied to 20 test cases available in test data and the predictions are submitted in appropiate format to the Course Project Prediction Quiz for grading.

#2. Summary 

Using latest technology we are able to collect a large amount of data about personal activity. These type of devices gives us quantified self movement data, for us to have a clear picture about our health and to find different behaviours in our bodies. One of the most common things is getting the number of exercises and not the quality of execution. We will use data from accelerometers of 6 participants. They performed barbell lifts correctly and incorrently in 5 different ways. More information : http://groupware.les.inf.puc-rio.br/har

#3. Data Preprocessing 

#3.1 Data
Train set : https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

Test set : https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

#3.2 Loading packages
```{r, echo=TRUE}
library(knitr)
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(corrplot)
library(ggplot2)
library(caret)
library(ggcorrplot)
library(GGally)
set.seed(1000)
```

#3.3 Data Preprocessing 
```{r, echo = FALSE}
TrainFile <- "C:/workspace/R training/Prediction-Assignment-Writeup/pml-training.csv"
TestFile <- "C:/workspace/R training/Prediction-Assignment-Writeup/pml-testing.csv"
```

```{r, echo=TRUE}
train <- read.csv(TrainFile)
dim(train)
test <- read.csv(TestFile)
dim(test)
```

```{r, echo=TRUE}
# creating data partition
inTrain <- createDataPartition(train$classe, p = 0.7, list = FALSE)
TrainSet <- train[inTrain,]
TestSet <- train[-inTrain,]

dim(TrainSet)
dim(TestSet)


#Both datasets have equal number of variables. Nor, for a better prediction we will exclude NA's, the Near Zone Variance and the Identification variables

#Removing variables with Nearly Zero Variance

nearzerovariance <- nearZeroVar(TrainSet)
TrainSet <- TrainSet[,-nearzerovariance]
TestSet <- TestSet[,-nearzerovariance]
dim(TrainSet)
dim(TestSet)

#Removing variables which are almost NA

RmvNA <- sapply(TrainSet, function(x) mean(is.na(x))) > 0.95
TrainSet <- TrainSet[,RmvNA==FALSE]
TestSet <- TestSet[,RmvNA==FALSE]
dim(TrainSet)
dim(TestSet)

#Removing Identification columns 1:5

TrainSet <- TrainSet[,-(1:5)]
TestSet <- TestSet[,-(1:5)]
dim(TestSet)
dim(TrainSet)


```

#4. Prediction Model Building

We will apply the most popular three solutions to model the regressions: Random Forest, Decision Tree and Generalized Boosted Model. A Confusion Matrix is plotted at the end of each model and the one one with the best accuracy will be used for the quiz predictions.

#4.1 Random Forest
```{r, echo=TRUE}
#model fitting
ctrlRF <- trainControl(method = "cv", number =3, verboseIter = FALSE)
modfitRF <- train(classe ~., data = TrainSet, method = "rf", trControl = ctrlRF)
modfitRF$finalModel

# predicting on test dataset
predictRF <- predict(modfitRF, newdata = TestSet)
CMRF <- confusionMatrix(predictRF,TestSet$classe)
CMRF
#plot matrix results
plot(CMRF$table, col = CMRF$byClass, main = paste("RF - Accuracy = ", round(CMRF$overall['Accuracy'],4)))
```

#4.2 Decision Tree
```{r, echo = TRUE}
#model fitting
modfitDT <- rpart(classe ~ ., data=TrainSet, method="class")
rpart.plot(modfitDT)

#predicting on test dataset
predictDT <- predict(modfitDT, newdata=TestSet, type="class")
CMDT <- confusionMatrix(predictDT, TestSet$classe)
CMDT
#plot matrix results
plot(CMDT$table, col = CMDT$byClass, 
     main = paste("Decision Tree - Accuracy =",
                  round(CMDT$overall['Accuracy'], 4)))
```

#4.3 Generalized Boosted Model
```{r, echo = TRUE}
#model fitting
ctrlGBM <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
modFitGBM  <- train(classe ~ ., data=TrainSet, method = "gbm",
                    trControl = ctrlGBM, verbose = FALSE)
modFitGBM$finalModel

# predicting on test dataset
predictGBM <- predict(modFitGBM, newdata=TestSet)
CMGBM <- confusionMatrix(predictGBM, TestSet$classe)
CMGBM
# plot matrix results
plot(CMGBM$table, col = CMGBM$byClass, 
     main = paste("GBM - Accuracy =", round(CMGBM$overall['Accuracy'], 4)))
```

#5. Aplying the selected Model to the Test Data

The accuracy of the three models used is:
Rf : 0.9971
DT : 0.7648
GBM : 0.9856

Due to this results I will use the Random Forest model to predict the 20 results (testing dataset)
```{r, echo= TRUE}
predictTEST <- predict(modfitRF, newdata=test)
predictTEST
```